---
title: "CDC Diabetes Health Indicators"
author: "Robert Gomez"
format: html
editor: visual
---

# Approach

## Introduction

Explore CDC BRFSS survey data on lifestyle behaviors collected from a random sample of the U.S. general population in 2015. I plan to use the R rio package to import the dataset from my 607 GitHub repository and then apply dplyr to restructure the data in preparation for exploratory analysis.

## Data Source

<https://archive.ics.uci.edu/dataset/891/cdc+diabetes+health+indicators>

## Loading packages

```{r}
pacman::p_load( rio,          # Importing data
                tidyverse,    # dplyr and ggplot2
                summarytools  # summarizing data
                )
```

## Import Data

```{r}
brfss2015 <- import("https://raw.githubusercontent.com/RGomez560/Data607/refs/heads/main/diabetes_012_health_indicators_BRFSS2015.csv")
```

## Transform Data

```{r}
brfss2015v2 <- brfss2015 %>%
  rename(DiabetesStatus = Diabetes_012) %>%
  select(DiabetesStatus, BMI) %>%
  mutate(
    DiabetesStatus = recode(
      DiabetesStatus,
      "0" = "No Diabetes",
      "1" = "Pre Diabetes",
      "2" = "Yes Diabetes"
    )
  )
```

## Explore Data

Data structure

```{r}
str(brfss2015v2)
```

Data frequencies

```{r}
brfss2015v2 %>%
  freq(round.digits = 2, cumul = FALSE, report.nas = FALSE)
```

Data frequencies visuals

```{r}
ggplot(brfss2015v2, aes(x = DiabetesStatus, fill = DiabetesStatus)) +
  geom_bar() +
  scale_fill_hue(c = 40) +
  theme(legend.position = "none") +
  labs(x = "Diabetes Status", y = "Count")
```
